---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rules

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![R-CMD-check](https://github.com/tidymodels/rules/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/tidymodels/rules/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/tidymodels/rules/branch/main/graph/badge.svg)](https://app.codecov.io/gh/tidymodels/rules?branch=main)
[![CRAN status](https://www.r-pkg.org/badges/version/rules)](https://cran.r-project.org/package=rules)
<!-- badges: end -->

## Introduction

rules is a [parsnip](https://parsnip.tidymodels.org/) extension package with model definitions for rule-based models, including:

 * cubist models that have discrete rule sets that contain linear models with an ensemble method similar to boosting
 * classification rules where a ruleset is derived from an initial tree fit
 * _rule-fit_ models that begin with rules extracted from a tree ensemble which are then added to a regularized linear or logistic regression. 

## Installation

You can install the released version of rules from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("rules")
```

Install the development version from GitHub with:

``` r
# install.packages("pak")
pak::pak("tidymodels/rules")
```

## Available Engines

The rules package provides engines for the models in the following table.

```{r}
#| echo: false
#| message: false
library(parsnip)

parsnip_models <- get_from_env("models") |>
  setNames() |>
  purrr::map_dfr(get_from_env, .id = "model")

library(rules)

rules_models <- get_from_env("models") |>
  setNames() |>
  purrr::map_dfr(get_from_env, .id = "model")

dplyr::anti_join(
  rules_models, parsnip_models, 
  by = c("model", "engine", "mode")
) |>
  knitr::kable()
```

## Contributing

This project is released with a [Contributor Code of Conduct](https://www.contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

- For questions and discussions about tidymodels packages, modeling, and machine learning, please [post on RStudio Community](https://community.rstudio.com/new-topic?category_id=15&tags=tidymodels,question).

- If you think you have encountered a bug, please [submit an issue](https://github.com/tidymodels/rules/issues).

- Either way, learn how to create and share a [reprex](https://reprex.tidyverse.org/articles/articles/learn-reprex.html) (a minimal, reproducible example), to clearly communicate about your code.

- Check out further details on [contributing guidelines for tidymodels packages](https://www.tidymodels.org/contribute/) and [how to get help](https://www.tidymodels.org/help/).
